Linearizability Proofs Based On Forward Simulations

### 2017-01-21 draftCAV/stack.tex
started explaining commit points

Pushing a value to the stack proceeds in several steps: adding a node with maximal timestamp, asking for a new timestamp (given by the value of the shared integer {\tt TS}), and updating the timestamp of the added node. Popping a value from the stack consists in traversing all the lists, finding the first element which doesn't represent a removed value (i.e., {\tt taken} is {\tt false}) in each list, and selecting the element with the maximal timestamp. A compare-and-swap (CAS) is used to set the {\tt taken} flag of this element to {\tt true}. The procedure restarts if the CAS fails.

>>>>

Pushing a value to the stack proceeds in several steps: adding a node with maximal timestamp in the list associated to the thread executing the push (given by the special variable {\tt myTID}), asking for a new timestamp (given by the value of the shared integer {\tt TS}), and updating the timestamp of the added node. Popping a value from the stack consists in traversing all the lists, finding the first element which doesn't represent a removed value (i.e., {\tt taken} is {\tt false}) in each list, and selecting the element with the maximal timestamp. A compare-and-swap (CAS) is used to set the {\tt taken} flag of this element to {\tt true}. The procedure restarts if the CAS fails.

\begin{figure}[t]
\centering
\includegraphics[width=11cm]{fig-timeStamp.pdf}
\caption{An execution of the Time-Stamped Stack.}
\label{fig:commit}
\end{figure}

The push operations don't have a fixed linearization point because adding a node to a list and updating its timestamp are not executed in a single atomic step. The nodes can be added in an order which is not consistent with the order between the timestamps assigned later in the execution. Also, the value added by a push that just added an element to a list can be popped before the value added by a completed push (since it has a maximal timestamp). The same holds for pop operations. The only reasonable choice for a linearization point is a successful CAS (that results in updating the field {\tt taken}).

TODO GIVE AN EXAMPLE TO SHOW THE ISSUE WITH POP LINEARIZATION POINTS, AND TO SHOW THAT COMMIT POINTS ARE RIGHT LIMITS FOR THE INTERVAL OF A LIN POINT, I.E., A POP WITH A LATER COMMIT POINT CAN BE LINEARIZED BEFORE. NEED ONLY 2 POPS IN AN EXECUTION
